`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

---
name: rvertnet
layout: post
title: Vertnet - getting vertebrate museum record data and a quick map
date: 2012-09-19
author: Scott Chamberlain
tags: 
- R
- ropensci
- open access
- data
- API
---

We (rOpenSci) started a repo to wrap the API for [VertNet](http://vertnet.org/index.php), an open access online database of vertebrate specimen records across many collection holders. Find the open source code [here](https://github.com/ropensci/rvertnet) - please contribute if you are so inclined.  We had a great Google Summer of Code student, [Vijay Barve](http://vijaybarve.wordpress.com/) contributing to the repo this summer, so it is getting close to being CRAN-able. 

Most of the functions in the repo get you the raw data, but there were no functions to visualize the data.  Since much of the data records of latitude and longitude data, maps are a natural visualization to use.  

What follows is a quick example of using the basic `vertmap` function.

### First, let's install `rvertnet`
```{r install }
# install_github('rvertnet', 'ropensci') # uncomment if not installed already
library(rvertnet)
```

### First, let's get some data using `vertoccurrence`
```{r vertoccurrence }
out <- vertoccurrence(q = "larva", num = 100) # get records on keyword 'larva', limit to 100
nrow(out) # how many rows?
```

### Now map it using `vertmap`.  This is a very basic function: it simply cleans up the input data.frame, removing rows without lat/long data, and providing warnings when the input data.frame is not in the correct format.  `vertmap` uses the `ggplot2` framework for the map.  If you want to make you own map please do so -  this is just a simple fxn to get you started if you want to take a quick look at the data. 
```{r fig.width=8, fig.height=6, vertmap }
vertmap(input = out) # make a map using vertmap
```

*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/sckott.github.io/blob/master/_drafts/2012-09-19-rvertnet.Rmd) - or [.md file](https://github.com/sckott/sckott.github.io/tree/master/_posts/2012-09-19-rvertnet.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/), and nice knitr highlighting/etc. in in [RStudio](http://rstudio.org/).