--- 
name: global-names-resolver
layout: post
title: Hitting the Global Names Resolver API
date: 2012-07-15
tags: 
- API
- ropensci
- taxonomic
- resolve
- taxize
- ropensci
- taxize
---

## Example of using the Global Names Resolver API to check species names

*********

There are a number of options for resolution of taxonomic names. The [Taxonomic Name Resolution Service (TNRS)](http://tnrs.iplantcollaborative.org/) comes to mind. There is a new service for taxonomic name resoultion called the [Global Names Resolver](http://resolver.globalnames.org/). They describe the service thusly "_Resolve lists of scientific names against known sources. This service parses incoming names, executes exact or fuzzy matching as required, and displays a confidence score for each match along with its identifier._". 

*********

## Load required packages

### Just uncomment the code to use.

```{r highlight=TRUE}
# If you don't have them already
# install.packages(c("RJSONIO","plyr","devtools"))
# require(devtools)
# install_github("taxize_","ropensci")
library(RJSONIO); library(plyr); library(taxize)
```

## Get the data sources available

### Get just id's and names of sources in a data.frame
```{r highlight=TRUE}
tail( gnr_datasources(todf=T) )
```

*********

### Give me the id for EOL (Encyclopedia of Life)
```{r highlight=TRUE}
out <- gnr_datasources(todf=T)
out[out$title == "EOL", "id"]
```

*********

### Fuzzy search for sources with the word "zoo"
```{r highlight=TRUE}
out <- gnr_datasources(todf=T)
outdf <- out[agrep("zoo", out$title, ignore.case=T), ]
outdf[1:2,]
```

## Resolve some names

### Search for _Helianthus annuus_ and _Homo sapiens_, return a data.frame
```{r highlight=TRUE}
gnr(names = c("Helianthus annuus", "Homo sapiens"), returndf = TRUE)[1:2,]
```

*********

### Search for the same species, with only using data source 12 (i.e., EOL)
```{r highlight=TRUE}
gnr(names = c("Helianthus annuus", "Homo sapiens"), data_source_ids="12", returndf = TRUE)
```


### That's it. Have fun! And put bugs/comments/etc. [here](https://github.com/ropensci/taxize_/issues).

*********

### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/), and nice knitr highlighting/etc. in in [RStudio](http://rstudio.org/).

*********

### I prepared the markdown for this post by:
```{r highlight=TRUE, eval=FALSE}
KnitPost <- function(input, base.url = "/") {
	require(knitr)
	opts_knit$set(base.url = base.url)
	fig.path <- paste0("img/", sub(".Rmd$", "", basename(input)), "/")
	opts_chunk$set(fig.path = fig.path)
	opts_chunk$set(fig.cap = "center")
	render_jekyll()
	knit(input, envir = parent.frame())
}
setwd("/path/to/_posts")
KnitPost("/path/to/postfile.Rmd")
```
from [jfisher](http://jfisher-usgs.github.com/r/2012/07/03/knitr-jekyll/).