`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

---
name: rgbif-newfxns
layout: post
title: GBIF biodiversity data from R - more functions
date: 2012-10-08
author: Scott Chamberlain
tags: 
- R
- open access
- data
- GBIF
- biodiversity
- macroecology
---

We have been working on an R package to get GBIF data from R, with the stable version available through CRAN [here](URL), and the development version available on GitHub [here](http://github.com/rgbif). 

We had a Google Summer of code stuent work on the package this summer - you can see his work on the package over at his GitHub page [here]().  We have added some new functionality since his work, and would like to show it off. 

### Lets install rgbif first.
```{r install } 
# install_github('rgbif', 'ropensci') # uncomment if not already installed
library(rgbif); library(plyr); library(XML); library(httr)
````

### Get taxonomic information on a specific taxon or taxa in GBIF by their taxon concept keys.
```{r taxonget } 
( keys <- taxonsearch(scientificname = 'Puma concolor') ) # many matches to this search
taxonget(keys[[1]]) # let's get the first one - the first row in the data.frame is the one we searched for (51780668)
````

### The `occurrencedensity` function was renamed to `densitylist` because it is in the `density` API service, not the `occurrence` API service.  You can use `densitylist` to get a data.frame of total occurrence counts by one-degree cell for a single taxon, country, dataset, data publisher or data network.  Just a quick reminder of what the function can do:
```{r densitylist }
head( densitylist(originisocountrycode = "CA") )
````

### Using a related function, `density_spplist`, you can get a species list by one-degree cell as well.
```{r fig.width=8, fig.height=6, density_spplist } 
# Get a species list by cell, choosing one at random
density_spplist(originisocountrycode = "CO", spplist = "random")[1:10]
# density_spplist(originisocountrycode = "CO", spplist = "r") # can abbreviate the `spplist` argument
 
# Get a species list by cell, choosing the one with the greatest no. of records
density_spplist(originisocountrycode = "CO", spplist = "great")[1:10] # great is abbreviated from `greatest`

# Can also get a data.frame with counts instead of the species list
density_spplist(originisocountrycode = "CO", spplist = "great", listcount = "counts")[1:10,]
````

### You can now map point results, from fxns `occurrencelist` and those from `densitylist`, which plots them as points or as tiles, respectively.  Point map, using output from occurrencelist.
```{r fig.width=8, fig.height=5, gbifmap1 } 
out <- occurrencelist(scientificname = 'Puma concolor', coordinatestatus = TRUE, maxresults = 100, latlongdf = T)
gbifmap(input = out) # make a map, plotting on world map
````

### Point map, using output from occurrencelist, with many species plotted as different colors
```{r fig.width=10, fig.height=5, gbifmap2 } 
splist <- c('Accipiter erythronemius', 'Junco hyemalis', 'Aix sponsa', 'Buteo regalis')
out <- lapply(splist, function(x) occurrencelist(x, coordinatestatus = T, maxresults = 100, latlongdf = T))
gbifmap(out)
```

### Tile map, using output from densitylist, using results in Canada only.
```{r fig.width=8, fig.height=5, gbifmap3 } 
out2 <- densitylist(originisocountrycode = "CA") # data for Canada
gbifmap(out2) # on world map
gbifmap(out2, "Canada") # on Canada map
````

*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/schamberlain.github.com/blob/master/_drafts/2012-10-08-rgbif-newfxns.Rmd) - or [.md file](https://github.com/SChamberlain/schamberlain.github.com/tree/master/_posts/2012-10-08-rgbif-newfxns.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).