`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

---
name: get-taxa-downstream
layout: post
title: Getting taxonomic names downstream
date: 2012-10-16
author: Scott Chamberlain
tags: 
- R
- open access
- data
- taxonomy
- ropensci
- ritis
- taxize
---

It can be a pain in the ass to get taxonomic names. For example, I sometimes need to get all the Class names for a set of species.  This is a relatively easy problem using the [ITIS API](http://www.itis.gov/ws_description.html) (example below).  

The much harder problem is getting all the taxonomic names downstream. ITIS doesn't provide an API method for this - well, they do ([`getHirerachyDownFromTSN`](http://www.itis.gov/ws_hierApiDescription.html#getHierarchyDn)), but it only provides direct children (e.g., the genera within a tribe - but it won't give all the species within each genus).  

So in the `taxize` package, we wrote a function called `downstream`, which allows you to get taxonomic names to any downstream point, e.g.:

+ get all Classes within Animalia,
+ get all Species within a Family
+ etc.

### Install packages.  You can get other packages from CRAN, but taxize is only on GitHub for now. 
```{r install } 
# install_github('ritis', 'ropensci') # uncomment if not already installed
# install_github('taxize_', 'ropensci') # uncomment if not already installed
library(ritis); library(taxize)
````

### Get upstream taxonomic names.
```{r upstream } 
# Search for a TSN by scientific name
df <- searchbyscientificname("Tardigrada")
tsn <- df[df$combinedname %in% "Tardigrada","tsn"]

# Get just one immediate higher taxonomic name
gethierarchyupfromtsn(tsn = tsn)

# Get full hierarchy upstream from TSN
getfullhierarchyfromtsn(tsn = tsn)
````

### Get taxonomc names downstream.
```{r downstream } 
# Get genera downstream fromthe Class Bangiophyceae
downstream(846509, "Genus")

# Get families downstream from __Acridoidea__
downstream(650497, "Family")

# Get species downstream from __Ursus__
downstream(180541, "Species")
````

*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/sckott.github.io/blob/master/_drafts/2012-10-16-get-taxa-downstream.Rmd) - or [.md file](https://github.com/sckott/sckott.github.io/tree/master/_posts/2012-10-16-get-taxa-downstream.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).