`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

---
name: rmetadata-questions
layout: post
title: Scholarly metadata from R - a bit more
date: 2012-10-01
author: Scott Chamberlain
tags: 
- R
- open access
- data
- metadata
- OAI-PMH
---

I did a recent post on getting scholarly metadata from the web via R [here](http://sckott.github.io/2012/09/rmetadata/). 

What kinds of things might you use rmetadata for?  In other words, what are somem use cases?

### Before we do anything else, lets install rmetadata first.
```{r install } 
# install_github('rmetadata', 'ropensci') # uncomment if not already instealled
library(rmetadata); library(plyr); library(XML); library(httr)
````

### One use case would be to ask how many records does each data holder have. 
```{r identify }
# Just one source
count_identifiers("pensoft")

# Many sources
ldply(c("datacite","Academic Commons","pensoft","arXiv"), function(x) count_identifiers(x, seconds=5))
````

### I also tried to get counts from all ~1800 OAI-PMH data providers. Note that I ran this on Amazon EC2 to speed things up.
```{r eval=FALSE, alldat }
# All of them, takes a while, run on AWS AMI if you can
out <- llply(providers[,2], function(x) count_identifiers(x, useurl = T), .inform=T)
outdf <- ldply(out)
str(outdf)
outdf_ <- outdf[!is.na(as.numeric(as.character(outdf$count))), ]  # just those with counts
```

Which comes up with very few data holders providing this number in their metadata.  

```{r alldatresults }
library(ggplot2)
dat <- read.csv("/Users/ScottMac/Downloads/oai_pmh_counts.csv")[,-1]
dat$count <- as.numeric(dat$count) # to numeric

# A distribution of values
ggplot(dat, aes(count)) + geom_density(fill="grey") + scale_x_log10() + theme_bw(base_size=18)

# so median of 3856 and mean of ~65000
median(dat$count)
mean(dat$count)
```

### One question could be to ask about frequency of use of a set of words in metadata records across different repositories.  We can start by getting a list of identifiers for some repositories using `md_listidentifiers`. Since there is a lot of identifiers in each repo, let's limit the dates to the last month basically.
```{r else }
# ids <- llply(dat[c(1:3,9:11,20,21,23,25,28,36:38,42,45:48,50),"provider"], function(x) md_listidentifiers(x, from = "2012-08-24"), .progress="text")
# names(ids) <- dat[c(1:3,9:11,20,21,23,25,28,36:38,42,45:48,50),"provider"] # rename each element
dat <- read.csv("/Users/ScottMac/Downloads/oai_pmh_counts.csv")[,-1]
dat$count <- as.numeric(dat$count) # to numeric
dat <- dat[-c(8,30,77,80,81,82,84,110,206),]
ids <- llply(dat[1:2,"provider"], function(x) md_listidentifiers(x, from = "2012-08-24"), .progress="text", .inform=T)
sapply(ids, function(x) length(x[[1]])) # lots of records :)
names(ids) <- dat[1:2,"provider"] # rename each element

# Get the records
md_getrecord("pensoft", "10.3897/zookeys.1.10") # a single record

registerDoMC(cores=4)
llply(ids[[1]][[1]][1:10], md_getrecord, .parallel=F) # all records from the ids vector

# plot em

```


*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/sckott.github.io/blob/master/_drafts/2012-09-21-rmetadata-questions.Rmd) - or [.md file](https://github.com/sckott/sckott.github.io/tree/master/_posts/2012-09-21-rmetadata-questions.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/), and nice knitr highlighting/etc. in in [RStudio](http://rstudio.org/).