`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

*********

## UPDATE: changed data source so that the entire example can be run by anone on their own machine. Also, per Joachim's suggestion, I put a box around the blown up area of the map. Thoughts?

## Here's a quick demo of creating a map with an inset within it using ggplot. The inset is achieved using the `gridExtra` package. 

*********

### Install libraries
```{r }
library(ggplot2); library(maps); library(maptools); library(gridExtra); library(rgeos)
```

*********

### Create a data frame
```{r }
dat <- data.frame(ecosystem = rep(c("oak","steppe", "prairie"), each=8),
									lat = rnorm(24, mean=51, sd=1), lon = rnorm(24, mean=-113, sd=5))
head(dat)
```

*********

### Get maps using the maps library
```{r }
# Get a map of Canada
canadamap <- data.frame(map("world", "Canada", plot=FALSE)[c("x","y")])

# Get a map of smaller extent
canadamapsmall <- canadamap[canadamap$x < -90 & canadamap$y < 54,]

# Make inset rectangle to show area of zoom
canadamapsmall_ <- na.omit(canadamapsmall) # omit NA's

	# This should get your corner points for the box, picking min and max of lat and lon
insetrect <- data.frame(xmin = min(canadamapsmall_$x), xmax = max(canadamapsmall_$x), 
												ymin = min(canadamapsmall_$y), ymax = max(canadamapsmall_$y))
insetrect
```

*********

### Make the maps
```{r tidy=FALSE}
# The inset map, all of Canada
a <- ggplot(canadamap) + 
	theme_bw(base_size = 22) +
	geom_path(data = canadamap, aes(x, y), colour = "black", fill = "white") +
	geom_rect(data = insetrect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), alpha=0, colour="blue", size = 1, linetype=1) +
	coord_map(projection = "mercator") +
	theme(panel.border = element_rect(colour = 'black', size = 1, linetype=1),
			 panel.grid.major = element_blank(), panel.grid.minor=element_blank(),
			 panel.background = element_rect( fill = 'white'),
			 legend.position = c(0.15,0.80), legend.key = element_blank(),
			 axis.ticks = element_blank(), axis.text.x = element_blank(),
			 axis.text.y = element_blank()) +
	labs(x = '', y = '')

# The larger map, zoomed in, with the data
b <- ggplot(dat, aes(lon, lat, colour=ecosystem)) +
	theme_bw(base_size = 22) +
	geom_jitter(size=4, alpha=0.6) +
	geom_path(data = canadamapsmall, aes(x, y), colour = "black", fill = "white") +
	scale_size(guide="none") +
	coord_map(projection = "mollweide") +
	theme(panel.border = element_rect(colour = 'black', size = 1, linetype=1),
			 panel.grid.major = element_blank(), panel.grid.minor=element_blank(),
			 panel.background = element_rect( fill = 'white'),
			 legend.position = c(0.1,0.20), legend.text = element_text(size=12, face='bold'), 
			 legend.title=element_text(size=12, face='bold'), legend.key = element_blank(),
			 axis.ticks = element_line(size = 2)) +
	labs(x = '', y = '')
```

*********

### Print the two maps together, one an inset of the other
#### This approach uses the `gridExtra` package for flexible alignment, etc. of ggplot graphs
```{r fig.width=8, fig.height=6}
grid.newpage()
vpb_ <- viewport(width = 1, height = 1, x = 0.5, y = 0.5) # the larger map
vpa_ <- viewport(width = 0.4, height = 0.4, x = 0.8, y = 0.8) # the inset in upper right
print(b, vp = vpb_)
print(a, vp = vpa_)
```

*********

### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/schamberlain.github.com/blob/master/_drafts/2012-08-22-ggplot-inset-map.Rmd).