`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

*********

I created an R package a while back to interact with some APIs that serve up data on what our elected represenatives are up to, including the [New York Times Congress API](http://developer.nytimes.com/), and the [Sunlight Labs API](http://services.sunlightlabs.com/).

What kinds of things can you do with `govdat`?  Here are a few examples. 

*********

### How do the two major parties differ in the use of certain words (searches the congressional record using the Sunlight Labs Capitol Words API)?
```{r}
# install_github('govdat', 'schamberlain')
library(govdat); library(reshape2); library(ggplot2)

dems <- sll_cw_dates(phrase='science', start_date='1996-01-20',
   end_date='2012-09-01', granularity='year', party='D', printdf=TRUE)
repubs <- sll_cw_dates(phrase='science', start_date='1996-01-20',
   end_date='2012-09-01', granularity='year', party='R', printdf=TRUE)
df <- melt(
	rbind(data.frame(party = rep("D",nrow(dems)), dems), data.frame(party = rep("R",nrow(repubs)),repubs))
	)
df$count <- as.numeric(df$count)

ggplot(df, aes(yearmonth, count, colour = party, group = party)) + 
	geom_line() + 
	scale_colour_manual(values=c("blue","red")) + 
	labs(y="use of the word 'Science'") +
	theme_bw(base_size=18) +
	opts(axis.text.x = theme_text(size = 10), 
			 panel.grid.major = theme_blank(), 
			 panel.grid.minor = theme_blank(), 
			 legend.position = c(0.2,0.8))
```

*********

### Let's get some data on donations to individual elected representatives.
```{r}
library(plyr)

# Let's get Nancy Pelosi's entity ID
sll_ts_aggregatesearch('Nancy Pelosi')[[1]]

# Her entity ID
sll_ts_aggregatesearch('Nancy Pelosi')[[1]]$id

# And search for her top donors by sector
nancy <- ldply(sll_ts_aggregatetopsectors(sll_ts_aggregatesearch('Nancy Pelosi')[[1]]$id))
nancy # but just abbreviations for sectors
data(sll_ts_sectors) # load sectors abbrevations data
nancy2 <- merge(nancy, sll_ts_sectors, by="sector") # attach full sector names
nancy2_melt <- melt(nancy2[,-1], id.vars=3)
nancy2_melt$value <- as.numeric(nancy2_melt$value)

# and lets plot some results
ggplot(nancy2_melt, aes(sector_name, value)) +
	geom_bar() +
	coord_flip() +
	facet_wrap(~ variable, scales="free", ncol=1)

## It looks like a lot of individual donations (the count facet) by finance/insurance/realestate, but by amount, the most (by slim margin) is from labor organizations.
```

*********

### Or we may want to get a bio of a congressperson. Here we get Todd Akin of MO. And some twitter searching too? Indeed.
```{r}
out <- nyt_cg_memberbioroles('A000358') # cool, lots of info, output cutoff for brevity
out[[3]][[1]][1:2]

# we can get her twitter id from this bio, and search twitter using twitteR package
akintwitter <- out[[3]][[1]]$twitter_id

# install.packages("twitteR")
library(twitteR)
tweets <- userTimeline(akintwitter, n=100)
tweets[1:5] # there's some gems in there no doubt
```

*********

### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/schamberlain.github.com/blob/master/_drafts/2012-09-01-gov-dat.Rmd) - or [.md file](https://github.com/SChamberlain/schamberlain.github.com/tree/master/_posts/2012-09-01-gov-dat.md).

*********

### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/), and nice knitr highlighting/etc. in in [RStudio](http://rstudio.org/).