`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

---
name: rgbif-newfxns
layout: post
title: GBIF biodiversity data from R - more functions
date: 2012-10-08
author: Scott Chamberlain
tags: 
- R
- open access
- data
- GBIF
- biodiversity
- macroecology
- ropensci
---

#### UPDATE: In response to Jarrett's query I laid out a separate use case in which you may want to query by higher taxonomic rankings than species. See below.  In addition, added examples of querying by location in reply to comments by seminym. 

*****

We have been working on an R package to get GBIF data from R, with the stable version available through CRAN [here](URL), and the development version available on GitHub [here](http://github.com/rgbif). 

We had a Google Summer of code stuent work on the package this summer - you can see his work on the package over at his GitHub page [here]().  We have added some new functionality since his work, and would like to show it off. 

### Lets install rgbif first.
```{r install } 
# install_github('rgbif', 'ropensci') # uncomment if not already installed
library(rgbif); library(plyr); library(XML); library(httr); library(maps); library(ggplot2)
````

### Get taxonomic information on a specific taxon or taxa in GBIF by their taxon concept keys.
```{r taxonget } 
( keys <- taxonsearch(scientificname = 'Puma concolor') ) # many matches to this search
taxonget(keys[[1]]) # let's get the first one - the first row in the data.frame is the one we searched for (51780668)
````

### The `occurrencedensity` function was renamed to `densitylist` because it is in the `density` API service, not the `occurrence` API service.  You can use `densitylist` to get a data.frame of total occurrence counts by one-degree cell for a single taxon, country, dataset, data publisher or data network.  Just a quick reminder of what the function can do:
```{r densitylist }
head( densitylist(originisocountrycode = "CA") )
````

### Using a related function, `density_spplist`, you can get a species list by one-degree cell as well.
```{r fig.width=8, fig.height=6, density_spplist } 
# Get a species list by cell, choosing one at random
density_spplist(originisocountrycode = "CO", spplist = "random")[1:10]
# density_spplist(originisocountrycode = "CO", spplist = "r") # can abbreviate the `spplist` argument
 
# Get a species list by cell, choosing the one with the greatest no. of records
density_spplist(originisocountrycode = "CO", spplist = "great")[1:10] # great is abbreviated from `greatest`

# Can also get a data.frame with counts instead of the species list
density_spplist(originisocountrycode = "CO", spplist = "great", listcount = "counts")[1:10,]
````

### You can now map point results, from fxns `occurrencelist` and those from `densitylist`, which plots them as points or as tiles, respectively.  Point map, using output from occurrencelist.
```{r fig.width=8, fig.height=5, gbifmap1 } 
out <- occurrencelist(scientificname = 'Puma concolor', coordinatestatus = TRUE, maxresults = 100, latlongdf = T)
gbifmap(input = out) # make a map, plotting on world map
````

### Point map, using output from occurrencelist, with many species plotted as different colors
```{r fig.width=10, fig.height=5, gbifmap2 } 
splist <- c('Accipiter erythronemius', 'Junco hyemalis', 'Aix sponsa', 'Buteo regalis')
out <- lapply(splist, function(x) occurrencelist(x, coordinatestatus = T, maxresults = 100, latlongdf = T))
gbifmap(out)
```

### Tile map, using output from densitylist, using results in Canada only.
```{r fig.width=8, fig.height=5, gbifmap3 } 
out2 <- densitylist(originisocountrycode = "CA") # data for Canada
gbifmap(out2) # on world map
gbifmap(out2, region="Canada") # on Canada map
```

*****

### We can also query by higher taxonomic rankings, and map all lower species within that ranking. Above we queried by scientificname, but we can also query by higher taxonomy. 7071443 is the taxonconceptkey for 'Bacillariophyceae', a Class which includes many lower species. 
```{r fig.width=8, fig.height=5, algae } 
out <- densitylist(taxonconceptKey=7071443)
gbifmap(out)
```

### seminym asked about querying by area. You can query by area, though slightly differently for occurrencelist and densitylist functions. For occurrencelist you can search using min and max lat and long values (and min an max altitude, pretty cool, eh).  
```{r fig.width=8, fig.height=5, byarea_occurr } 
# Get occurrences or density by area, using min/max lat/long coordinates
out <- occurrencelist(minlatitude=30, maxlatitude=35, minlongitude=-100, maxlongitude=-95, 
							 coordinatestatus = T, maxresults = 5000, latlongdf = T)

# Using `geom_point`
gbifmap(out, "state", "texas", geom_point)

# Using geom_jitter to move the points apart from one another
gbifmap(out, "state", "texas", geom_jitter, position_jitter(width = 0.3, height = 0.3))

# And move points a lot
gbifmap(out, "state", "texas", geom_jitter, position_jitter(width = 1, height = 1))
```

### And you can query by area in `densitylist` by specifying a place using the `originisocountrycode` argument (as done in an above example).  Just showing the head of the data.frame here.  
```{r fig.width=8, fig.height=5, byarea_density } 
# Get density by place, note that you can't use the lat and long arguments in densitylist
head( densitylist(originisocountrycode="CA") )
```

*********
#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/sckott.github.io/blob/master/_drafts/2012-10-08-rgbif-newfxns.Rmd) - or [.md file](https://github.com/sckott/sckott.github.io/tree/master/_posts/2012-10-08-rgbif-newfxns.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).