opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, cache=FALSE)

---
name: is-invasive
layout: post
title: Is invasive?
date: 2012-12-13
author: Scott Chamberlain
tags: 
- R
- ropensci
- taxize
- invasive
---

The Global Invasive Species Database (GISD) (see their website for more info [here](http://www.issg.org/database/welcome/)) has data on the invasiveness status of many species. From `taxize` you can now query the GISD database. 

Introducing the function `gisd_isinvasive`. This function was contributed to `taxize` by [Ignasi Bartomeus](http://www.bartomeus.cat/es/ignasi/), a postdoc at the Swedish University Agricultural Sciences. 

There are two possible outputs from using `gisd_isinvasive`: "Invasive" or "Not in GISD". If you use `simplify=TRUE` in the function you get "Invasive" or "Not in GISD", but if you use `simplify=FALSE` you get verbose description of the invasive species instead of just "Invasive" (and you still just get "Not in GISD"). 

***************

### Install taxize from GitHub
```{r }
# install_github('taxize_', 'ropensci') # install if you don't already have the GitHub version
library(taxize)
```

***************

### Make a vector of species
```{r }
sp <- c("Carpobrotus edulis", "Rosmarinus officinalis", "Nasua nasua", "Martes melampus", "Centaurea solstitialis")
```

### Using the function `gisd_isinvasive` you can query the GISD database for the invasiveness status of your species, at least according to GISD. Calling `gisd_isinvasive` with the second parameter set to default `simplify=FALSE`, you get verbose output, with details on the species. 
```{r }
gisd_isinvasive(sp)
```

### Simpler output, just the invasive status. 
```{r }
gisd_isinvasive(sp, simplify = TRUE)
````

*********

#### Get the .Rmd file used to create this post [at my github account](https://github.com/sckott/scott/blob/gh-pages/_drafts/2012-12-13-is-invasive.Rmd) - or [.md file](https://github.com/sckott/scott/blob/gh-pages/_posts/2012-12-13-is-invasive.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).