`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

*********

## So I was trying to figure out a fast way to make matrices with randomly allocated 0 or 1 in each cell of the matrix. I reached out on Twitter, and got many responses (thanks tweeps!). 

*********

### Here is the solution I came up with. See if you can tell why it would be slow.
```{r}
mm <- matrix(0, 10, 5) 
apply(mm, c(1,2), function(x) sample(c(0,1),1))
```

*********

### Ted Hart (@distribecology) replied first with:
```{r}
matrix(rbinom(10*5, 1, 0.5), ncol = 5, nrow = 10)
```

*********


### Next, David Smith (@revodavid) and Rafael Maia (@hylospar) came up with about the same solution. 
```{r}
m <- 10
n <- 5
matrix(sample(0:1, m*n, replace = TRUE), m, n)
```

*********


### Then there was the solution by Luis Apiolaza (@zentree).
```{r}
m <- 10
n <- 5
round(matrix(runif(m*n), m, n))
```

*********

### Last, a solution was proposed using `RcppArmadillo`, but I couldn't get it to work on my machine, but here is the function anyway if someone can. 
```{r eval=FALSE}
library(inline); library(RcppArmadillo)
f <- cxxfunction(body='return wrap(arma::randu(5,10));', plugin='RcppArmadillo')
```

*********

### And here is the comparison of system.time for each solution. 
```{r }
mm <- matrix(0, 10, 5)
m <- 10
n <- 5

system.time( replicate(1000, apply(mm, c(1,2), function(x) sample(c(0,1),1)) ) ) # @recology_
system.time( replicate(1000, matrix(rbinom(10*5,1,.5),ncol=5,nrow=10) ) ) # @distribecology
system.time( replicate(1000, matrix(sample(0:1, m*n, replace=TRUE),m,n) ) ) # @revodavid & @hylospar
system.time( replicate(1000, round(matrix(runif(m*n), m, n)), ) ) # @zentree
```

### If you want to take the time to learn C++ or already know it, the RcppArmadillo option would likely be the fastest, but I think (IMO) for many scientists, especially ecologists, we probably don't already know C++, so will stick to the next fastest options. 

*********

### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/schamberlain.github.com/blob/master/_drafts/2012-08-30-making-matrices.Rmd).