opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, cache=FALSE)

---
name: ropensci-collaboration
layout: post
title: Visualizing rOpenSci collaboration
date: 2013-02-15
author: Scott Chamberlain
tags: 
- R
- ropensci
- github
- API
---

We ([rOpenSci](http://ropensci.org/)) have been writing code now for our R packages for a couple years, so perhaps it is time to take a look back at the data. What data you ask? The commits data from GitHub, or the data that records who did what and when. 

Using the [Github commits API](http://developer.github.com/v3/repos/commits/) we can gather data on who commited code to a Github code repo, and when they did it. Then we can visualize this hitorical record. 

***************

```{r eval=FALSE}

```

***************

In addition, there are quite a few people that have committed code now to rOpenSci repos, calling for a network vizualization of course. 

```{r}

```

***************

#### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/scott/blob/gh-pages/_drafts/2013-02-15-ropensci-collaboration.Rmd) - or [.md file](https://github.com/SChamberlain/scott/blob/gh-pages/_posts/2013-02-15-ropensci-collaboration.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).