opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, cache=FALSE)

```{r  echo=FALSE, warning=FALSE	}
require(knitcitations); cleanbib()
````

---
name: open-science-challenge
layout: post
title: Open Science Challenge
date: 2013-01-08
author: Scott Chamberlain
tags: 
- R
- ropensci
---

### The setup

Science is becoming more open in many areas: publishing, data sharing, lab notebooks, and software. There are many benefits to open science. For example, sharing research data alongside your publications leads to increased citation rate `r citep("10.1371/journal.pone.0000308")`. In addition, data is becoming easier to share and reuse thanks to efforts like [FigShare](http://figshare.com/) and [Dryad](http://datadryad.org/). 

If you don't understand the problem we are currently facing due to lack of open science, just watch this video:

<iframe width="560" height="315" src="http://www.youtube.com/embed/N2zK3sAtr-4" frameborder="0" allowfullscreen></iframe>

***************

### The challenge

To speed up adoption of open science approaches, we ([ropensci](http://ropensci.org/)) have just kicked off the [rOpenSci Open Science Challenge](http://ropensci.org/open-science-challenge/). If you aren't familiar with rOpenSci, it is a software collective connecting scientists to open science data on the web. Since R is the most popular programming language for life scientists, it made sense to do this in R (instead of Python e.g.). 

***************

### But what is it about

What is the challenge about? We want to facilitate open science. Since we write software, this challenge asks you to propose a project using one or more of our packages - or perhaps you want to propose a new dataset to connect to R. The rOpenSci core developer team will help you with any problems using our packages, and attempt to modify packages according to feedback from participants. Do you use one or more of our R packages? If you do, great. If not, check out our packages [here](http://ropensci.org/packages/index.html). 

***************

### __How to apply?__ 

Just send an email to [info@ropensci.org](mailto:info@ropensci.org?subject=rOpenSci Open Science Challenge). 

***************

### __The deadline?__ 

January 31, 2013

***************

#### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/scott/blob/gh-pages/_drafts/2013-01-08-open-science-challenge.Rmd) - or [.md file](https://github.com/SChamberlain/scott/blob/gh-pages/_posts/2013-01-08-open-science-challenge.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/), and [knitcitations](https://github.com/cboettig/knitcitations) from [Carl Boettiger](http://www.carlboettiger.info/).

***************

#### References
```{r results="asis", echo=FALSE }
bibliography(style="html")
````