opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, cache=FALSE)

---
name: tnrs-use-case
layout: post
title: Resolving species names when you have a lot of them
date: 2013-01-25
author: Scott Chamberlain
tags: 
- R
- ropensci
- taxize
- taxonomic
---

### __taxize use case: Resolving species names when you have a lot of them__

Species names can be a pain in the ass, especially if you are an ecologist. We ecologists aren't trained in taxonomy, yet we often end up with huge species lists.  Of course we want to correct any spelling errors in the names, and get the newest names for our species, resolve any synonyms, etc. 

We are building tools into our R package [`taxize`](http://ropensci.github.com/taxize_/), that will let you check your species names to make sure they are correct. 

An important use case is when you have a lot of species. Someone wrote to us recently, saying that they had thousands of species, and they wanted to know how to check their species names efficiently in R. 

Below is an example of how to do this. 

***************

#### Install taxize
```{r}
# install_github("taxize_", "ropensci") # install the GitHub version, not the CRAN version, uncomment if you don't have it installed
library(taxize)
```

***************

#### Get some species, in this case all species in the Scrophulariaceae family from theplantlist.org
```{r}
tpl_get(dir_ = "~/foo2", family = "Scrophulariaceae")
dat <- read.csv("~/foo2/Scrophulariaceae.csv")
```

***************

#### Lets grab the species and concatenate to genus_species
```{r}
species <- as.character(ddply(dat[,c("Genus","Species")], .(), transform, gen_sp = as.factor(paste(Genus,Species,sep=" ")))[,4])
```

***************

#### It's better to do many smaller calls to a web API instead of few big ones to be nice to the database maintainers.
```{r}
## Define function to split up your species list into useable chuncks
slice <- function(input, by=2){
	starts <- seq(1,length(input),by)
	tt <- lapply(starts, function(y) input[y:(y+(by-1))])
	llply(tt, function(x) x[!is.na(x)])
}
species_split <- slice(species, by=100)
```

***************

#### Query for your large species list with pauses between calls, with 3 seconds in between calls to not hit the web service too hard. Using POST method here instead of GET - required when you have a lot of species.
```{r}
tnrs_safe <- failwith(NULL, tnrs) # in case some calls fail, will continue
out <- llply(species_split, function(x) tnrs_safe(x, getpost="POST", sleep=3))

# Looks like we got some data back for each element of our species list
lapply(out, head)[1:2] # just look at the first two

# Now we can put them back together as so into one data.frame if you like
outdf <- ldply(out)
head(outdf)
```

Note that there are multiple names for some species because data sources have different names for the same species (resulting in more than one row in the data.frame 'outdf' for a species). We are leaving this up to the user to decide which to use. For example, for the species _Buddleja montana_ there are two names for in the output
```{r}
data <- ddply(outdf, .(submittedName), summarize, length(submittedName))
outdf[outdf$submittedName %in% as.character(data[data$..1 > 1,][6,"submittedName"]),]
```

The source iPlant matched the name, but NCBI actually gave back a genus of cricket (follow the link under the column uri for _Montana_). If you look at the page for _Buddleja_ on NCBI [here](http://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi?id=26473) there is no _Buddleja montana_ at all. 

Another thing we could do is look at the score that is returned. Let's look at those that are less than 1 (i.e., )
```{r}
outdf[outdf$score < 1,]
```

As we got this speies list from [theplantlist.org](http://www.theplantlist.org/), there aren't that many mistakes, but if it was my species list you know there would be many :)


***************

### That's it.  Try it out and let us know if you have any questions at [info@ropensci.org](mailto:info@ropensci.org), or [ask questions/report problems at GitHub](https://github.com/ropensci/taxize_/issues).

***************

#### Get the .Rmd file used to create this post [at my github account](https://github.com/SChamberlain/scott/blob/gh-pages/_drafts/2013-01-25-tnrs-use-case.Rmd) - or [.md file](https://github.com/SChamberlain/scott/blob/gh-pages/_posts/2013-01-25-tnrs-use-case.md).

#### Written in [Markdown](http://daringfireball.net/projects/markdown/), with help from [knitr](http://yihui.name/knitr/).