---
name: sofa-reboot
layout: post
title: sofa - reboot
date: 2014-11-18
author: Scott Chamberlain
tags:
- R
- API
- database
- couchdb
- nosql
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#>",
  warning=FALSE, 
  message=FALSE,
  fig.path="../public/img/2014-11-18-sofa/",
  fig.cap = ""
)
```

I've reworked `sofa` recently after someone reported a bug in the package. Since the last post on this package on 2013-06-21, there's a bunch of changes:

* Removed the `sofa_` prefix from all functions as it wasn't really necessary. 
* Replaced `rjson`/`RJSONIO` with `jsonlite` for JSON I/O.
* New functions:
  * `revisions()` - to get the revision numbers for a document.
  * `uuids()` - get any number of UUIDs - e.g., if you want to set document IDs with UUIDs
* Most functions that deal with documents are prefixed with `doc_`
* Simplified all code, reducing duplication
* All functions take `cushion` as the first parameter, for consistency sake.
* Changed `cushion()` function so that you can only register one cushion with each function call, 
and the function takes parameters for each element now, `name` (name of the cushion, whatever you want), `user` (user name, if applicable), `pwd` (password, if applicable), `type` (one of localhost, cloudant, or iriscouch), and `port` (if applicable).
* Changed package license from `CC0` to `MIT`

There's still more to do, but I'm pretty happy with the recent changes, and I hope at least some find the package useful. Also, would love people to try it out as all bugs are shallow and all that...

The following are a few examples of package use.

## Install CouchDB

Instructions [here](http://wiki.apache.org/couchdb/Installation)

## Start CouchDB

In your terminal 

```sh
couchdb
```

You can interact with your CouchDB databases as well in your browser. Navigate to [http://localhost:5984/_utils](http://localhost:5984/_utils)

## Install sofa

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("sckott/sofa")
```

```{r}
library('sofa')
```

## Authenticate - Cushions

As an example, here's how I set up details for connecting to my Cloudant couch:

```{r}
cushion(name = 'cloudant', user = '<user name>', pwd = '<password>', type = "cloudant")
```

By default there is a built-in `cushion` for localhost so you don't have to do that, unless you want to change those details, e.g., the port number. Right now cushions aren't preserved across R sessions, but working on that.

For example, I'll lay down a cushion for Cloudant, then I can call `cushions()` to see my cushions:

```{r eval=FALSE}
cushion(name = 'cloudant', user = '<user name>', pwd = '<pwd>', type = "cloudant")
cushions()
```

By default, if you don't provide a cushion name, you are using localhost.

## Ping the server

```{r}
ping()
```

Nice, it's working.

## Create a new database, and list available databases

```{r echo=FALSE}
if(is.null(dbinfo(dbname = "sofadb")$error)) deletedb(dbname='sofadb')
```

```{r}
createdb(dbname='sofadb')
```

see if its there now

```{r}
listdbs()
```

## Create documents

Create a document WITH a name (uses PUT)

```{r}
doc1 <- '{"name":"sofa","beer":"IPA"}'
doc_create(dbname="sofadb", doc=doc1, docid="a_beer")
```

Create a document WITHOUT a name (uses POST)

```{r}
doc2 <- '{"name":"sofa","icecream":"rocky road"}'
doc_create(dbname="sofadb", doc=doc2)
```

## List documents

List them

```{r}
alldocs(dbname="sofadb")
```

Optionally include the documents, returned as a list by default as it would be hard to parse an endless number of document formats. 

```{r}
alldocs(dbname="sofadb", include_docs = TRUE)
```

## Update a document

Change _IPA_ (india pale ale) to _IPL_ (india pale lager). We need to get revisions first as we need to include revision number when we update a document.

```{r}
(revs <- revisions(dbname = "sofadb", docid = "a_beer"))
```

```{r}
newdoc <- '{"name":"sofa","beer":"IPL"}'
doc_update(dbname = "sofadb", doc = newdoc, docid = "a_beer", rev = revs[1])
```

Should be two revisions now

```{r}
revisions(dbname = "sofadb", docid = "a_beer")
```

## Get headers for a document

```{r}
doc_head(dbname = "sofadb", docid = "a_beer")
```

## JSON vs. list

Across all/most functions you can request json or list as output with the `as` parameter. 

```{r}
listdbs(as = "list")
```

```{r}
listdbs(as = "json")
```

## Curl options

Across all functions you can pass in curl options. We're using `httr` internally, so you can use `httr` helper functions to make some curl options easier. Examples:

Verbose output

```{r}
library("httr")
listdbs(config=verbose())
```

Progress

```{r}
listdbs(config=progress())
```

Set a timeout

```{r eval=FALSE}
listdbs(config=timeout(seconds = 0.001))
#> 
#> Error in function (type, msg, asError = TRUE)  : 
#>    Operation timed out after 3 milliseconds with 0 out of -1 bytes received
```


## Full text search

I'm working on an R client for Elaticsearch called `elastic` - find it at [https://github.com/ropensci/elastic](https://github.com/ropensci/elastic)

Thinking about where to include functions to allow `elastic` and `sofa` to work together...if you have any thoughts hit up the issues. I'll probably include helper functions for CouchDB search in the `elastic` package, interfacing with the [CouchDB plugin for Elasticsearch](https://github.com/elasticsearch/elasticsearch-river-couchdb).
