---
name: analogsea-v01
layout: post
title: analogsea - v0.1 notes
date: 2014-06-18
author: Scott Chamberlain
tags:
- R
- API
- cloud
---


My [last blog ](http://recology.info/2014/05/analogsea/) post introduced the R package I'm working on `analogsea`, an R client for the Digital Ocean API. 

Things have changed a bit, including fillig out more functions for all API endpoints, and incorparting feedback from Hadley and Karthik. The package is as `v0.1` now, so I thought I'd say a few things about how it works. 

Note that Digital Ocean's v2 API is in beta stage now, so the current version of `analogsea` at `v0.1` works with their v1 API. The [v2 branch of analogsea](https://github.com/sckott/analogsea/tree/v2) is being developed for their v2 API.

First, installation

Note: I did try to submit to CRAN, but Ripley complained about the package name so I'd rather not waste my time esp since people using this likely will already know about `install_github()`.

```{r eval=FALSE}
devtools::install_github("sckott/analogsea")
```

Load the library

```{r}
library("analogsea")
```

Authenticate has changed a bit. Whereas auth details were stored as environment variables before, I'm just using R's options. `do_auth()` will ask for your Digital Ocean details. You can enter them each R session, or store them in your `.Rprofile` file. After successful authentication, each function simply looks for your auth details with `getOption()`.  You don't have to use this function first, though if you don't your first call to another function will ask for auth details. 

```{r results='hide'}
do_auth()
```

`sizes`, `images`, and `keys` functions have changed a bit, by default outputting a `data.frame` now. 

List available regions

```{r }
regions()
```

List available sizes

```{r}
sizes()
```

List available images

```{r}
head(images())
```

List ssh keys

```{r}
keys()
```

One change that's of interest is that most of the various `droplets_*()` functions take in the outputs of other `droplets_*()` functions. This means that we can pipe outputs of one `droplets_*()` function to another, including non-`droplet_*` functions (see examples). 

Let's create a droplet:

```{r create1, eval=FALSE}
(res <- droplets_new(name="foo", size_slug = '512mb', image_slug = 'ubuntu-14-04-x64', region_slug = 'sfo1', ssh_key_ids = 89103))
```

```r
$droplet
$droplet$id
[1] 1880805

$droplet$name
[1] "foo"

$droplet$image_id
[1] 3240036

$droplet$size_id
[1] 66

$droplet$event_id
[1] 26711810
```

List my droplets

This function used to be `do_droplets_get()`

```{r}
droplets()
```

As mentioned above we can now pipe output of `droplet*()` functions to other `droplet*()` functions. 

Here, pipe output of lising droplets `droplets()` to the `events()` function

```{r}
droplets() %>% events()
```

In this case there were no event ids to get event data on. 

Here, we'll get details for the droplet we just created, then pipe that to `droplets_power_off()`

```{r}
droplets(1880805) %>% droplets_power_off
```

Then pipe it again to `droplets_power_on()`

```{r echo=FALSE}
Sys.sleep(12)
```

```{r}
droplets(1880805) %>% 
  droplets_power_on
Sys.sleep(6)
droplets(1880805)$droplets$status
```

Why not use more pipes?

```{r, tidy=FALSE}
droplets(1880805) %>% 
  droplets_power_off %>%
  droplets_power_on %>%
  events
```

-------

Last time I talked about installing R, RStudio, etc. on a droplet. I'm still working out bugs in that stuff, but do test out so it can get better faster. See `do_install()`.