---
name: taxize-workflows
layout: post
title: taxize workflows
date: 2014-12-02
author: Scott Chamberlain
tags:
- R
- API
- taxize
- taxonomy
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#>",
  warning=FALSE,
  message=FALSE
)
```

A missed chat on the rOpenSci website the other day asked:

> Hi there, i am trying to use the taxize package and have a .csv file of species names to run through taxize updating them. What would be the code i would need to run to achieve this?

One way to answer this is to talk about the basic approach to importing data, doing stuff to the data, then recombining data. There are many ways to do this, but I'll go over a few of them.

## Install taxize

```{r eval=FALSE}
install.packages("taxize")
install.packages("downloader")
```

```{r}
library("taxize")
```

## Import data

We'll use Winston Chang's new `downloader` package to avoid problems with `https`, and get a dataset from our ropensci datasets repo [https://github.com/ropensci/datasets](https://github.com/ropensci/datasets)

```{r}
downloader::download("https://raw.githubusercontent.com/ropensci/datasets/master/planttraits/morphological.csv", "morphological.csv")
dat <- read.csv("morphological.csv", stringsAsFactors = FALSE)
head(dat)
```

After importing data, there are a variety of approaches you could take:

1. Vector: Take species names as vector from your `data.frame`, cleaning them, then re-attching to the `data.frame` later, or
2. In-Place: Use for loops or `lapply` family functions to iterate over each name while simultaneously re-inserting into the `data.frame`

## 1. Vector

Make a vector of names

```{r}
splist <- dat$species
```

Then proceed to do name cleaning, e.g, we can use the `tnrs` function to see if any names are potentially not spelled correctly. 

```{r}
tnrs_out <- tnrs(splist, source = "iPlant_TNRS")
head(tnrs_out)
```

Those with score of less than 1 may have misspellings

```{r}
tnrs_out[ tnrs_out$score < 1, ]
```

So let's take the `acceptedname` column as a the new names and assign to a new vector

```{r}
cleaned_names <- tnrs_out$acceptedname
```

Then join names back, replacing them, or adding as a new column

Replace

```{r}
dat$species <- cleaned_names
head(dat)
```

New column

```{r}
dat$species_cleaned <- cleaned_names
head(dat)
```

## 2. In-place

You can use functions from the `dplyr` package to `split-apply-combine`, where `split` is split apart your vector for each taxon, `apply` to apply a function or functions to do name cleaning, then `combine` to put them back together. 

Here, we'll attach taxonomic ids from the Catalogue of Life to each species (each row) (with just a subset of the data to save time):

```{r}
library("dplyr")
tbl_df(dat)[1:5,] %>%
  rowwise() %>%
  mutate(colid = get_colid(species)) %>%
  select(species, colid)
```

Let's do something a bit more complicated. Get common names for each taxon in a new column, if more than 1, concatenate into a single character string for easy inclusion in a `data.frame`

```{r}
sci2comm_concat <- function(x){
  temp <- sci2comm(x, db = "eol")
  if(length(temp) == 0) NA else paste0(temp[[1]], collapse = ", ")
}

dat_new <- tbl_df(dat)[1:5,] %>%
  rowwise() %>%
  mutate(comm = sci2comm_concat(species))
```

To see the new column, do 

```{r}
dat_new %>% select(comm)
```
