---
name: cowsay
layout: post
title: cowsay - ascii messages and warnings for R
date: 2014-02-20
author: Scott Chamberlain
tags:
- r
- API
- ascii
---

## The history

Cowsay is a terminal program that generates ascii pictures of a cow saying what you tell the cow to say in a bubble. See the Wikipedia page for more information: [http://en.wikipedia.org/wiki/Cowsay](http://en.wikipedia.org/wiki/Cowsay). 

Install cowsay to use in your terminal (on OSX):

```
brew update
brew install cowsay
```

Type `cowsay hello world!`, and you get:

```
 ______________
< hello world! >
 --------------
        \   ^__^
         \  (oo)\_______
            (__)\       )\/\
                ||----w |
                ||     ||
```

Optionally, you can install [fortune](http://en.wikipedia.org/wiki/Fortune_(Unix)) to get pseudorandom messages from a database of quotations. On OSX do `brew install fortune`, then you can pipe a fortune quote to `cowsay`:

```
fortune | cowsay
```

And get something like:

```
 ______________________________________
/ "To take a significant step forward, \
| you must make a series of finite     |
| improvements." -- Donald J. Atwood,  |
\ General Motors                       /
 --------------------------------------
        \   ^__^
         \  (oo)\_______
            (__)\       )\/\
                ||----w |
                ||     ||
```

You can also get different animals. Try `cowsay -f tux <yourmessage>`

## Cowsay in R

Why cowsay for R?  Why not. You never know what you will learn in fun side projects. Basically, this cowsay R package we are making prints messages that you pass to the function `say`.  There are three arguments to the `say` function:

* __what__: What do you want to say?  You can pass it a custom message, anything you want, like _what's up_, or _howdy!_. You can also get R's version of fortunes, quotes about R. Just pass the exact term _forture_. If you want a fact about cats from the [Cat Facts API](http://catfacts-api.appspot.com/), pass in _catfact_. Last, you can get the current time by passing _time_ to this parameter.
* __by__: Type of animal, one of cow, chicken, poop, cat, ant, pumpkin, ghost, spider, rabbit, pig, snowman, or frog. If you want more animals, send a pull request, or ask and at some point it will be added. 
* __type__: One of message (default), warning, or string (returns string). You could use string to pass into other functions, etc., instead of printing a warning or message.

There are three other contributors so far (a big thanks to them):

* Tyler Rinker
* Thomas Leeper
* Noam Ross

### Installation

```{r eval=FALSE}
library(devtools)
install_github("cowsay", "sckott")
```

```{r}
library(cowsay)
```

p.s. or `install_github("sckott/cowsay")` if you have a newer version of devtools

### Get time

```{r comment=NA}
say('time')
```

```{r comment=NA}
say("time", "chicken")
```

### Vary type of output, default calls message

```{r comment=NA}
say("hello world")
```

```{r comment=NA}
say("hello world", type="warning")
```

```{r comment=NA}
say("hello world", type="string")
```

### Catfacts!!!!

From the [catfacts API](http://catfacts-api.appspot.com/), we can get random cat facts. If you put in _catfact_ you by default get a cat saying it. 

```{r comment=NA}
say("catfact", "cat")
```

### R fortunes

```{r comment=NA}
say("fortune")
```

```{r comment=NA}
say("fortune", "pig")
```

### Incorporate into a function

Define a function

```{r}
foo <- function(x){
  if(x<5) say("woops, x should be 5 or greater")
}
```

Call the function, with an error on purpose

```{r comment=NA}
foo(3)
```

Or capture a warning or message and pass to the `say` function

```{r}
foo2 <- function(x){
  err <- tryCatch(x^2, error=function(e) e)
  say(err$message, "frog")
}
```

Then call the function 

```{r comment=NA}
foo2("hello")
```

Awesome. Much better to have an error message from a frog than just the harsh console alone :)