---
name: rplos-pubs-country
layout: post
title: Publications by author country
date: 2014-12-03
author: Scott Chamberlain
tags:
- R
- API
- publications
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

I just missed another chat on the rOpenSci website:

> I want to know the number of publications by people from a certain country, but I dont know how to achieve this...

Fun! Let's do that. It's a bit complicated because there is no field like geography of the authors. But there are affiliation fields, from which we can collect data we need.

## Installation

You'll need the GitHub version for the coutry names data, or just use the CRAN version, and get country names elsewhere. 

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rplos")
```

```{r}
library("rplos")
```

## Get the data

```{r}
articles <- searchplos(q='*:*', limit = 5,
    fl=c("id","author_affiliate"), 
    fq=list('article_type:"Research Article"', "doc_type:full"))
```

## Search for country names in affilitation field

```{r}
(countries <- lapply(articles$data$author_affiliate, function(x){
  out <- sapply(isocodes$name, function(z) grepl(z, x))
  isocodes$name[out]
}))
```

You can combine this data with the previously collected data:

```{r}
# Helper function
splitem <- function(x){
  if(length(x) == 0) { NA } else {
    if(length(x) > 1) paste0(x, collapse = ", ") else x
  }
}

articles$data$countries <- sapply(countries, splitem)
head(articles$data)
```

## Bigger data set

Okay, cool, lets do it on a bigger data set, and this time, we'll get another variable `counter_total_all`, which is the combination of page views/pdf downloads for each article. This will allow us to ask _Is number of countries included in the authors related to page views?_. I have no idea if this question makes sense, but nonetheless, it is a question :)

```{r}
articles <- searchplos(q='*:*', limit = 1000,
    fl=c("id","counter_total_all","author_affiliate"), 
    fq=list('article_type:"Research Article"', "doc_type:full"))
```

Get countries

```{r}
countries <- lapply(articles$data$author_affiliate, function(x){
  out <- sapply(isocodes$name, function(z) grepl(z, x))
  isocodes$name[out]
})
df <- articles$data
df$countries <- sapply(countries, splitem)
```

Let's remove those rows with 0 countries, since the authors must be from somewhere, so the country name matching must have errored. 

```{r}
df$n_countries <- sapply(countries, length)
df <- df[ df$n_countries > 0, ]
```

Plot data

```{r}
library("ggplot2")
ggplot(df, aes(n_countries, as.numeric(counter_total_all))) +
  geom_point() +
  labs(y="total page views") + 
  theme_grey(base_size = 16)
```

Conclusion: meh, maybe, maybe not

## Into rplos

We'll probably add a function like this into `rplos`, as a convenient way to handle this use case.
