---
name: ckanr-intro
layout: post
title: Intro to alpha ckanr - R client for CKAN RESTful API 
date: 2014-11-26T11:42:36Z
author: Scott Chamberlain
tags:
- R
- API
- github
- database
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#>",
  warning=FALSE, 
  message=FALSE
)
```

Recently I had need to create a client for scraping museum metadata to help out some folks that use that kind of data. It's called [musemeta](https://github.com/ropensci/musemeta). One of the data sources in that package uses the open source _data portal software_ [CKAN](http://ckan.org/), and so we can interact with [the CKAN API](http://docs.ckan.org/en/latest/api/index.html) to get data. Since many groups can use CKAN API/etc infrastucture because it's open source, I thought why not have a general purpose R client for this, since [there are other clients](https://github.com/ckan/ckan/wiki/CKAN-API-Clients) for Python, PHP, Ruby, etc. 

Here's a bit of an intro:

## Setup

Get/load packages

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/ckanr")
```

```{r}
library("ckanr")
```

> Note: the default URL is for http://data.techno-science.ca/. You can change that in the `url` parameter

## Some package details

All API alls are `POST` requests, and handled through a helper function called `ckan_POST()`. See `ckanr:::ckan_POST` to see the function.

There are some common parameters across most functions that are worth describing a bit:

* `offset` (numeric) Where to start getting items from (optional, default: 0)
* `limit` (numeric) The maximum number of items to return (optional, default: 31)
* `url` Base url to use. Default: [http://data.techno-science.ca](http://data.techno-science.ca)
* `as` (character) One of `list` (default), `table`, or `json`. Parsing with table option uses `jsonlite::fromJSON(..., simplifyDataFrame = TRUE)`, which attempts to parse data to `data.frame`'s when possible, so the result can vary.
* `...` Curl args passed on to `httr::POST`

## Changes

```{r}
changes(limit = 2, as = "table")
```

## List datasets

```{r}
datasets(as = "table")
```

## List tags

```{r}
tag_list('aviation', as='table')
```

## Show tags

Subset for readme brevity

```{r}
tag_show('Aviation')$packages[[1]][1:3]  
```

## List groups

```{r}
group_list(as='table')
```

## Show groups

Subset for readme brevity

```{r}
group_show('communications', as='table')$users
```

## Show a package

```{r}
package_show('34d60b13-1fd5-430e-b0ec-c8bc7f4841cf', as='table')$resources
```

## Search for packages

```{r}
out <- package_search(q = '*:*', rows = 2, as="table")$results
out[, !names(out) %in% 'resources']
```

## Search for resources

```{r}
resource_search(q = 'name:data', limit = 2, as='table')
```

## Future work

* There's already [an issue](https://github.com/ropensci/ckanr/issues/4) to add support for [DataStore](http://docs.ckan.org/en/latest/maintaining/datastore.html)
* This client needs to be tested against many other CKAN API instances to make sure it's robust
* Add a test suite
* Use cases: would be nice to include in the package documentation use cases
* Other things? Get in touch on twitter `@recology_` or below
