---
name: rsunlight
layout: post
title: rsunlight - v0.3 released
date: 2014-08-11
author: Scott Chamberlain
tags:
- R
- API
- government
---

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(fig.path="../public/img/2014-08-11-rsunlight/")
```

My [last blog post on this package](http://recology.info/2014/05/rsunlight/) was so long ago the package wrapped both New York Times APIs and Sunlight Labs APIs and the package was called `govdat`. I split that package up into `rsunlight` for Sunlight Labs APIs and `rtimes` for some New York Times APIs. `rtimes` is [in development at Github](https://github.com/ropengov/rtimes).

We've updated the package to include four sets of functions, one set for each of four Sunlight Labs APIs (with a separate prefix for each API):

* Congress API (`cg_`)
* Open States API (`os_`)
* Capitol Words API (`cw_`)
* Influence Explorer API (`ie_`)

Then there are many methods for each API.

## rsunlight intro

### Installation

First, installation

```{r eval=FALSE}
devtools::install_github("ropengov/rsunlight")
```

Load the library

```{r}
library("rsunlight")
```

### Congress API

Search for Fed level bills that include the term _health care_ in them.

```{r}
res <- cg_bills(query='health care')
head(res$results[,1:4])
```

Search for bills that have the two terms _transparency_ and _accountability_ within 5 words of each other in the bill.

```{r}
res <- cg_bills(query='transparency accountability'~5)
head(res$results[,1:4])
```

### Open States API

Search State Bills, in this case search for the term _agriculture_ in Texas.

```{r}
res <- os_billsearch(terms = 'agriculture', state = 'tx')
head(res)
```

Search for legislators in California (_ca_) and in the democratic party

```{r}
res <- os_legislatorsearch(state = 'ca', party = 'democratic', fields = c('full_name','+capitol_office.phone'))
head(res)
```

Now you can call each representative, yay!

### Capitol Words API

Search for phrase _climate change_ used by politicians between September 5th and 16th, 2011:

```{r}
head(cw_text(phrase='climate change', start_date='2011-09-05', end_date='2011-09-16', party='D')[,c('speaker_last','origin_url')])
```

Plot mentions of the term _climate change_ over time for Democrats vs. Republicans

```{r fig.width=10, fig.height=6} 
library('ggplot2')
dat_d <- cw_timeseries(phrase='climate change', party="D")
dat_d$party <- rep("D", nrow(dat_d))
dat_r <- cw_timeseries(phrase='climate change', party="R")
dat_r$party <- rep("R", nrow(dat_r))
dat_both <- rbind(dat_d, dat_r)
ggplot(dat_both, aes(day, count, colour=party)) +
   geom_line() +
   theme_grey(base_size=20) +
   scale_colour_manual(values=c("blue","red"))
```

### Influence Explorer API

Search for contributions of equal to or more than `$20,000,000`.

```{r}
ie_contr(amount='>|20000000')[,c('amount','recipient_name','contributor_name')]
```

Top industries, by contributions given. _UNKOWN_ is a very influential industry. Of course _law firms_ are high up there, as well as _real estate_. I'm sure _oil and gas_ is embarrased that they're contributing less than _pulic sector unions_. 

```{r fig.width=10, fig.height=6}
(res <- ie_industries(method='top_ind', limit=10))
res$amount <- as.numeric(res$amount)
ggplot(res, aes(reorder(name, amount), amount)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels=dollar) +
  theme_grey(base_size = 14) +
  labs(x='Industry', y='Amount')
```

-------

## Feedback

Please do use `rsunlight`, and let us know what you want fixed, new features, etc.

## Still to come:

* Functions to visualize data from each API. You can do this yourself, but a few functions will be created to help those that are new to R.
* Vectorize functions so that you can give many inputs to a function instead of a single input.
* test suite: embarrasingly, there is no test suite yet, boo me.
* I plan to push `rsunlight` to CRAN soon as `v0.3`
