---
name: museum-aamsf
layout: post
title: Museum metadata - the Asian Art Museum of San Francisco
date: 2014-12-10
author: Scott Chamberlain
tags:
- R
- history
- scraping
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
warning = FALSE,
message = FALSE
)
```

I was in San Francisco last week for an altmetrics conference at PLOS. While there, I visited the [Asian Art Museum](http://www.asianart.org/), just the [Roads of Arabia exhibition](http://www.asianart.org/exhibitions_index/roads-of-arabia).

It was a great exhibit. While I was looking at the pieces, I read many labels, and thought, "hey, what if someone wants this metadata"? 

Since we have an R package in development for scraping museum metadata (called [musemeta](https://github.com/ropensci/musemeta)), I just started some scraping code for this museum. Unfortunately, I don't think the pieces from the Roads of Arabia exhibit are on their site, so no metadata to get. But they do have their main collection searchable online at [http://www.asianart.org/collections/collection](http://www.asianart.org/collections/collection). Examples follow. 

## Installation

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/musemeta")
```

```{r}
library("musemeta")
```

## Get metadata for a single object

You have to get the ID for the piece from their website, e.g., `11462` from the url `http://searchcollection.asianart.org/view/objects/asitem/nid/11462`. Once you have an ID you can pass it in ot the `aam()` function.

```{r}
(out <- aam(11462))
```

The output is printed for clarity, but you can dig into each element, like 

```{r}
out$label
```

## Get metadata for many objects

The `aam()` function is not vectorized, but you can easily get data for many IDs via `lapply` type functions, etc. 

```{r}
lapply(c(17150,17140,17144), aam)
```

## No search, boo

Note that there is no search functionality yet for this source. Maybe someone can add that via pull requests :)

## Like the others

The others sources in `musemeta` mostly work the same way as the above.
