---
name: geojson-io
layout: post
title: geojsonio - a new package to do geojson things
date: 2015-04-30
author: Scott Chamberlain
sourceslug: _drafts/2015-04-30-geojson-io.Rmd
tags:
- R
- geojson
- topojson
- geospatial
---

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`geojsonio` converts geographic data to GeoJSON and TopoJSON formats - though the focus is mostly on GeoJSON 

For those not familiar GeoJSON it is a format for encoding a variety of geographic data structures. GeoJSON supports the following geometry types: Point, LineString, Polygon, MultiPoint, MultiLineString, MultiPolygon, and GeometryCollection. These geometry types are also found in [well known text (WKT)](http://en.wikipedia.org/wiki/Well-known_text), and have equivalents in R's spatial classes. Read the [spec](http://geojson.org/geojson-spec.html) for more detailed information. 

Other great geojson resources:

* GeoJSON lint - lint your geojson - [http://geojsonlint.com/](http://geojsonlint.com/)
* GeoJSON.io - make maps with geojson input or draw maps and get geojson - [http://geojson.io/](http://geojson.io/)

Functions in this package are organized first around what you're working with or want to get, geojson or topojson, then convert to or read from various formats:

* `geojson_list()` - convert to GeoJSON as R list format
* `geojson_json()` - convert to GeoJSON as json
* `geojson_read()`/`topojson_read()` - read a GeoJSON/TopoJSON file from file path or URL
* `geojson_write()` - write a GeoJSON file locally (no write TopoJSON yet)

Each of the above functions have methods for various objects/classes, including `numeric`, `data.frame`, `list`, `SpatialPolygons`, `SpatialLines`, `SpatialPoints`, etc. (including the classes in `rgeos`)

Use cases for this package include (but not limited to, obs.) the following:

* Get data in GeoJSON json format, and you want to get it into a list in R.
* Get data into GeoJSON format to use downstream to make a interactive map
  * in R (e.g., with [leaflet](https://github.com/rstudio/leaflet))
  * or in another context (e.g., using javascript with mapbox/leaflet)
* Data is in a data.frame/matrix/list and you want to make GeoJSON format data.
* Data is in one of the many spatial classes (e.g., `SpatialPoints`) and you want GeoJSON
* You need to add styling to your data - can do with this package for certain data types.
* You want to check that your GeoJSON data is valid - two ways to do it in geojsonio.
* Combine objects together (e.g., a point and a line), either from two `geo_list` objects, or two `json` objects. See `?geojson-add`

## Install

See the github repo for notes about dependencies [https://github.com/ropensci/geojsonio#install](https://github.com/ropensci/geojsonio#install). 

CRAN version or the dev version from GitHub

```{r eval=FALSE}
install.packages("geojsonio")
devtools::install_github("sckott/geojsonio")
```

```{r}
library("geojsonio")
```

## GeoJSON

### Convert various formats to geojson

From a `numeric` vector of length 2

as _json_

```{r}
geojson_json(c(32.45, -99.74))
```

as a __list__

```{r output.lines=1:10}
geojson_list(c(32.45, -99.74))
```

From a `data.frame`

as __json__

```{r}
geojson_json(us_cities[1:2, ], lat = 'lat', lon = 'long')
```

as a __list__

```{r output.lines=1:10}
geojson_list(us_cities[1:2, ], lat = 'lat', lon = 'long')
```

From `SpatialPolygons` class

```{r output.lines=1:10}
library('sp')
poly1 <- Polygons(list(Polygon(cbind(c(-100, -90, -85, -100),
  c(40, 50, 45, 40)))), "1")
poly2 <- Polygons(list(Polygon(cbind(c(-90, -80, -75, -90),
  c(30, 40, 35, 30)))), "2")
(sp_poly <- SpatialPolygons(list(poly1, poly2), 1:2))
```

to __json__

```{r}
geojson_json(sp_poly)
```

to a __list__

```{r output.lines=1:10}
geojson_list(sp_poly)
```

From `SpatialPoints` class

```{r}
x <- c(1, 2, 3, 4, 5)
y <- c(3, 2, 5, 1, 4)
(s <- SpatialPoints(cbind(x, y)))
```

to __json__

```{r}
geojson_json(s)
```

to a __list__

```{r output.lines=1:10}
geojson_list(s)
```

### Combine objects

`geo_list` + `geo_list`

> Note: `geo_list` is the output type from `geojson_list()`, it's just a list with a class attached so we know it's geojson :)

```{r output.lines=1:10}
vec <- c(-99.74, 32.45)
a <- geojson_list(vec)
vecs <- list(c(100.0, 0.0), c(101.0, 0.0), c(100.0, 0.0))
b <- geojson_list(vecs, geometry = "polygon")
a + b
```

`json` + `json`

```{r}
c <- geojson_json(c(-99.74, 32.45))
vecs <- list(c(100.0, 0.0), c(101.0, 0.0), c(101.0, 1.0), c(100.0, 1.0), c(100.0, 0.0))
d <- geojson_json(vecs, geometry = "polygon")
c + d
```

### Write geojson

```{r}
geojson_write(us_cities[1:2, ], lat = 'lat', lon = 'long')
```

## Topojson

In the current version of this package you can read topojson. Writing topojson was in this package, but is gone for now - will come back later as in interface to [topojson](https://github.com/mbostock/topojson) via [V8](https://github.com/jeroenooms/V8).

Read from a file

```{r eval=FALSE}
file <- system.file("examples", "us_states.topojson", package = "geojsonio")
out <- topojson_read(file)
```

Read from a URL

```{r eval=FALSE}
url <- "https://raw.githubusercontent.com/shawnbot/d3-cartogram/master/data/us-states.topojson"
out <- topojson_read(url)
```

## Lint geojson

There are two ways to do this in this package. 

### lint, locally

Uses the javascript library [geojsonhint](https://github.com/mapbox/geojsonhint) from Mapbox. We're running this locally via the [V8](http://cran.rstudio.com/web/packages/V8/) package.

Good

```{r}
lint('{"type": "Point", "coordinates": [-100, 80]}')
```

Bad

```{r}
lint('{"type": "Rhombus", "coordinates": [[1, 2], [3, 4], [5, 6]]}')
```

### validate, with a web service

Uses the web service at [http://geojsonlint.com/](http://geojsonlint.com/)

Good

```{r}
validate('{"type": "Point", "coordinates": [-100, 80]}')
```

Bad

```{r}
validate('{"type": "Rhombus", "coordinates": [[1, 2], [3, 4], [5, 6]]}')
```

## To do

* I'd like to replace `rgdal` with javascript libraries to read from various file types (kml, shp, etc.) and convert to geojson. This is [in development](https://github.com/ropensci/geojsonio/tree/js), and will come in the next version of this package most likely. This should make installation a bit easier as we won't have to depend on `rgdal` and `GDAL`
* Performance improvements. Some operations already use the gdal or geos C libraries, so are quite fast, though the round trip to disk and back does take significant time. I'd like to speed this up. 
* More input types. We already have operations (json, list, etc.) for lots of input types (data.frame, list, sp classes), but likely there will be more added. 
* Most likely add functions `topojson_list()`, `topojson_json()`
