---
name: couch-dataframes
layout: post
title: PUT dataframes on your couch
date: 2015-03-12
author: Scott Chamberlain
sourceslug: _drafts/2015-03-12-couch-dataframes.Rmd
tags:
- R
- couchdb
- sofa
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE
)
```

It would be nice to easily push each row or column of a data.frame into CouchDB instead of having to prepare them yourself into JSON, then push in to couch. I recently added ability to push data.frame's into couch using the normal `PUT /{db}` method, and added support for the couch bulk API.

## Install

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("sckott/sofa")
```

```{r}
library("sofa")
```

## PUT /db

You can write directly from a data.frame, either by rows or columns. First, rows:

```{r echo=FALSE}
if(is.null(db_info(dbname = "mtcarsdb")$error)) db_delete(dbname='mtcarsdb')
```

Create a database

```{r}
db_create(dbname="mtcarsdb")
```

```{r}
out <- doc_create(mtcars, dbname="mtcarsdb", how="rows")
out[1:2]
```

Then by columns

```{r}
out <- doc_create(mtcars, dbname="mtcarsdb", how="columns")
out[1:2]
```

## Bulk API

The bulk API will/should be faster for larger data.frames

```{r echo=FALSE}
if(is.null(db_info(dbname = "bulktest")$error)) db_delete(dbname='bulktest')
```

We'll use part of the diamonds dataset

```{r}
library("ggplot2")
dat <- diamonds[1:20000,]
```

Create a database

```{r}
db_create(dbname="bulktest")
```

Load by row (could instead do each column, see `how` parameter), printing the time it takes

```{r}
system.time(out <- bulk_create(dat, dbname="bulktest"))
```

The returned data is the same as with `doc_create()`

```{r}
out[1:2]
```

So that's 20,000 rows in not that much time, not bad.

### not dataframes

You can also pass in lists or vectors of json as character strings, e.g., 

_lists_

```{r echo=FALSE}
if(is.null(db_info(dbname = "bulkfromlist")$error)) db_delete(dbname='bulkfromlist')
```

```{r}
row.names(mtcars) <- NULL # get rid of row.names
lst <- parse_df(mtcars, tojson=FALSE)
db_create(dbname="bulkfromlist")
out <- bulk_create(lst, dbname="bulkfromlist")
out[1:2]
```

_json_

```{r echo=FALSE}
if(is.null(db_info(dbname = "bulkfromchr")$error)) db_delete(dbname='bulkfromchr')
```

```{r}
strs <- as.character(parse_df(mtcars, "columns"))
db_create(dbname="bulkfromchr")
out <- bulk_create(strs, dbname="bulkfromchr")
out[1:2]
```
