---
name: rerddap
layout: post
title: rerddap - General purpose R client for ERDDAP servers
date: 2015-06-24
author: Scott Chamberlain
sourceslug: _drafts/2015-06-23-rerddap.Rmd
tags:
- R
- API
- NOAA
- web services
---

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines) == 1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1]) > 1) more else NULL,
            x[lines],
            if (length(x) > lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "../public/img/2015-06-24-rerddap/"
)
```

ERDDAP is a server built on top of OPenDAP - used by NOAA to serve many kinds of data. You can get gridded data ([griddap](http://upwell.pfeg.noaa.gov/erddap/griddap/documentation.html)), which lets you query from gridded datasets, or table data ([tabledap](http://upwell.pfeg.noaa.gov/erddap/tabledap/documentation.html)) which lets you query from tabular datasets. In terms of how we interface with them, there are similarties, but some differences too. We try to make a similar interface to both data types in `rerddap`.

ERDDAP is a powerful tool - in a world of heterogeneous data, it's often hard to combine data and serve it through the same interface, with tools for querying/filtering/subsetting the data. That is exactly what ERDDAP does. There are not no similarities, each often has lat/long data and usually a time component, but other variables in each dataset vary widely. 

## netcdf

`rerddap` supports netCDF format, and is the default when using the `griddap()` function. We use `ncdf` by default, but you can choose to use `ncdf4` instead.

## Caching

Data files downloaded are cached in a single hidden directory `~/.rerddap` on your machine. It's hidden so that you don't accidentally delete the data, but you can still easily delete the data if you like.

When you use `griddap()` or `tabledap()` functions, we construct a MD5 hash from the base URL, and any query parameters - this way each query is separately cached. Once we have the hash, we look in `~/.rerddap` for a matching hash. If there's a match we use that file on disk - if no match, we make a http request for the data to the ERDDAP server you specify.

## ERDDAP servers

You can get a data.frame of ERDDAP servers using the function `servers()`. Most I think serve some kind of NOAA data, but there are a few that aren't NOAA data. Here are a few:

```{r}
head(servers())
```


## Install

From CRAN

```{r eval=FALSE}
install.packages("rerddap")
```

Or development version from GitHub

```{r eval=FALSE}
devtools::install_github("ropensci/rerddap")
```

```{r}
library('rerddap')
```

## Search

First, you likely want to search for data, specifying whether to search for either `griddadp` or `tabledap` datasets. The default is `griddap`. 

```{r}
ed_search(query = 'size', which = "table")
```

```{r}
ed_search(query = 'size', which = "grid")
```

This gives back dataset titles and identifiers - with which you should be able to get a sense for which dataset you may want to fetch. 

## Information

After searching you can get more information on a single dataset

```{r output.lines=1:10}
info('whoi_62d0_9d64_c8ff')
```

Which is a simple S3 list but prints out pretty, so it's easy to quickly scan the printed output and see what you need to see to proceed. That is, in the next step you want to get the dataset, and you'll want to specify your search using some combination of values for latitude, longitude, and time. 

## griddap (gridded) data

First, get information on a dataset to see time range, lat/long range, and variables.

```{r}
(out <- info('noaa_esrl_027d_0fb5_5d38'))
```

Then query for gridded data using the `griddap()` function

```{r}
(res <- griddap(out,
  time = c('2012-01-01', '2012-01-30'),
  latitude = c(21, 10),
  longitude = c(-80, -70)
))
```

The output of `griddap()` is a list that you can explore further. Get the summary

```{r}
res$summary
```

Or get the dimension variables (just the names of the variables for brevity here)

```{r}
names(res$summary$dim)
```

Get the data.frame (beware: you may want to just look at the `head` of the data.frame if large)

```{r}
res$data
```

You can actually still explore the original netcdf summary object, e.g., 

```{r}
res$summary$dim$time
```

## tabledap (tabular) data

`tabledap` is data that is not gridded by lat/lon/time. In addition, the query interface is a bit different. Notice that you can do less than, more than, equal to type queries, but they are specified as character strings. 

```{r output.lines=1:10}
(out <- info('erdCalCOFIfshsiz'))
```

```{r}
(dat <- tabledap(out, 'time>=2001-07-07', 'time<=2001-07-10', 
                 fields = c('longitude', 'latitude', 'fish_size', 'itis_tsn', 'scientific_name')))
```

Since both `griddap()` and `tabledap()` give back data.frame's, it's easy to do downstream manipulation. For example, we can use `dplyr` to filter, summarize, group, and sort:

```{r}
library("dplyr")
dat$fish_size <- as.numeric(dat$fish_size)
df <- tbl_df(dat) %>% 
  filter(fish_size > 30) %>% 
  group_by(scientific_name) %>% 
  summarise(mean_size = mean(fish_size)) %>% 
  arrange(desc(mean_size))
df
```

Then make a cute little plot

```{r tidy=FALSE, fig.width=10}
library("ggplot2")
ggplot(df, aes(reorder(scientific_name, mean_size), mean_size)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  theme_grey(base_size = 20) +
  labs(y = "Mean Size", x = "Species")
```
