---
name: binomen
layout: post
title: binomen - taxonomic classes and parsing
date: 2015-01-19
author: Scott Chamberlain
sourceslug: _drafts/2015-01-19-binomen.Rmd
tags:
- R
- taxonomy
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

I maintain, along with other [awesome people](https://github.com/ropensci/taxize/graphs/contributors), the [taxize](https://github.com/ropensci/taxize) R package - a taxonomic toolbelt for R, for interacting with taxonomic data sources on the web. 

Taxonomy data is not standardized, but there are a lot of common elements, and there is a finite list of taxonomic ranks, and finite number of major taxonomic data sets. Thus, I've been interested in attempting to define a pseudo standard for expressing taxonomic data in R. The conversation [started a while back](https://github.com/ropensci/taxize/issues/261) in a GitHub issue, and hasn't moved very far. 

I decided to start playing with this more, which is easier to do in a separate pacakge. Thus: `binomen`. It's an attempt to define a set of taxonomic classes/objects in R, along with a suite of functions to help construct and parse these objects.

Would love any/all feedback. 

Here's some examples:

## Install

Install `binomen` 

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/binomen")
```

```{r}
library("binomen")
```

## Make a taxon

Make a taxon object

```{r}
(obj <- make_taxon(genus="Poa", epithet="annua", authority="L.",
                   family='Poaceae', clazz='Poales', 
                   kingdom='Plantae', variety='annua'))
```

Index to various parts of the object

The binomial

```{r}
obj$binomial
```

The authority

```{r}
obj$binomial$authority
```

The classification

```{r}
obj$classification
```

The family

```{r}
obj$classification$family
```

## Subset taxon objects

Get a single rank

```{r}
obj %>% select(family)
```

Get a range of ranks

```{r}
obj %>% range(kingdom, family)
```

Extract classification as a `data.frame`

```{r}
gethier(obj)
```

## Taxonomic data.frame's

Make one

```{r}
df <- data.frame(
  order=c('Asterales','Asterales','Fagales','Poales','Poales','Poales'),
  family=c('Asteraceae','Asteraceae','Fagaceae','Poaceae','Poaceae','Poaceae'),
  genus=c('Helianthus','Helianthus','Quercus','Poa','Festuca','Holodiscus'),
  stringsAsFactors = FALSE)
(df2 <- taxon_df(df))
```

Parse - get rank order matching _Fagales_

```{r}
df2 %>% select(order, Fagales)
```

get rank family matching _Asteraceae_

```{r}
df2 %>% select(family, Asteraceae)
```

get rank genus matching _Poa_

```{r}
df2 %>% select(genus, Poa)
```
