---
name: marine-regions
layout: post
title: Marine Regions data in R
date: 2015-12-16
author: Scott Chamberlain
sourceslug: _drafts/2015-12-16-marine-regions.Rmd
tags:
- R
- shp
- spatial
- geospatial
- marine
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

I've been at a hackathon focused on OBIS data this week in Belgium. One project idea was to make it easier to get at data based on one or more marine regions. I was told that Marineregions.org is most often used for shape files to get different regions to then do other work with.

I hacked together a package `mregions`, at [sckott/mregions][mregions]. Here I'll show how to get different marine regions, then pass those outputs directly to another package(s) to query for OBIS species occurrence data, then make maps.

## Install

```{r eval=FALSE}
devtools::install_github("sckott/mregions")
```

```{r}
library("mregions")
```

## Get list of regions

```{r}
"xxx"
```

## Download region - geojson

```{r}
"xxx"
```

## Download region - shp

```{r}
"xxx"
```

## Get regions, then OBIS data

```{r}
"xxx"
```

## Map data

```{r}
"xxx"
```

[mregions]: https://github.com/sckott/mregions
