---
name: the-new-way
layout: post
title: the new way - httsnap
date: 2015-04-29
author: Scott Chamberlain
sourceslug: _drafts/2015-04-29-the-new-way.Rmd
tags:
- R
- http
- httr
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

Inspired by `httpie`, a Python command line client as a sort of drop in replacement for `curl`, I am playing around with something similar-ish in R, at least in spirit. I started a little R pkg called `httsnap` with the following ideas:

* The web is increasingly a JSON world, so set `content-type` and `accept` headers to `applications/json` by default 
* The workflow follows logically, or at least should, from, _hey, I got this url_, to _i need to add some options_, to _execute request_
* Whenever possible, transform output to data.frame's - facilitating downstream manipulation via `dplyr`, etc.
* Do `GET` requests by default. Specify a different type if you don't want `GET`. Some functionality does GET by default, though in some cases you need to specify GET
* You can use non-standard evaluation to easily pass in query parameters without worrying about `&`'s, URL escaping, etc. (see `Query()`)
* Same for body params (see `Body()`)

## Install

Install and load `httsnap`

```{r eval=FALSE}
devtools::install_github("sckott/httsnap")
```

```{r}
library("httsnap")
library("dplyr")
```

## Functions so far

* `Get` - GET request
* `Query` - add query parameters
* `Authenticate` - add authentication details
* `Progress` - add progress bar
* `Timeout` - add a timeout
* `User_agent` - add a user agent
* `Verbose` - give verbose output
* `Body` - add a body
* `h` - add headers by key-value pair

These are named to avoid conflict with `httr`

## Intro

A simple `GET` request

```{r}
"http://httpbin.org/get" %>%
  Get()
```

You'll notice that `Get()` doesn't just get the response, but also checks for whether it was a good response (the HTTP status code), and extracts the data. 

Or you can just pass the URL into the function itself

```{r}
Get("http://httpbin.org/get")
```

You can buid up options by calling functions via pipes, and see what the options look like

```{r message=TRUE}
"http://httpbin.org/get" %>%
  Progress() %>%
  Verbose()
```

Then execute the GET request when you're ready

```{r message=TRUE}
"http://httpbin.org/get" %>%
  Progress() %>%
  Verbose() %>%
  Get()
```

## Example 1

Get scholarly article metadata from the Crossref API

```{r message=TRUE}
"http://api.crossref.org/works" %>%
  Query(query = "ecology") %>% 
  .$message %>% 
  .$items %>% 
  select(DOI, title, publisher)
```

## Example 2

Get Public Library of Science article metadata via their API, make a histogram of number of tweets for each article

```{r message=TRUE}
"http://api.plos.org/search" %>%
  Query(q = "*:*", wt = "json", fl = "id,journal,alm_twitterCount", rows = 100, 
        fq = 'alm_twitterCount:[100 TO 10000]') %>% 
  .$response %>% 
  .$docs %>% 
  .$alm_twitterCount %>% 
  hist()
```

## Notes

Okay, so this isn't drastically different from what `httr` already does, but its early days. 
