---
name: oai-client
layout: post
title: oai - an OAI-PMH client
date: 2015-09-11
author: Scott Chamberlain
sourceslug: _drafts/2015-09-11-oai-client.Rmd
tags:
- literature
- metadata
- datasets
- R
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`oai` is a general purpose client to work with any 'OAI-PMH' service. The 'OAI-PMH' protocol is described at [http://www.openarchives.org/OAI/openarchivesprotocol.html](http://www.openarchives.org/OAI/openarchivesprotocol.html). The main functions follow the OAI-PMH verbs: 

* `GetRecord`
* `Identify`
* `ListIdentifiers`
* `ListMetadataFormats`
* `ListRecords`
* `ListSets`

The repo is at [https://github.com/sckott/oai](https://github.com/sckott/oai)

I will be using this in a number of packages I maintain that use OAI-PMH data services. If you try it, let me know what you think. 

This package is heading to rOpenSci soon: [https://github.com/ropensci/onboarding/issues/19](https://github.com/ropensci/onboarding/issues/19)

Here's a few usage examples:

## Install

Is on CRAN now, but binaries may not be available yet. 

```{r eval=FALSE}
install.packages("oai")
```

Or install development version from GitHub

```{r eval=FALSE}
devtools::install_github("sckott/oai")
```

Load `oai`

```{r}
library("oai")
```

## Identify

```{r}
id("http://oai.datacite.org/oai")
```

## ListIdentifiers

```{r}
list_identifiers(from = '2011-05-01T', until = '2011-09-01T')
```

## Count Identifiers

```{r}
count_identifiers()
```

## ListRecords

```{r}
list_records(from = '2011-05-01T', until = '2011-08-15T')
```

## GetRecords

```{r}
get_records(c("oai:oai.datacite.org:32255", "oai:oai.datacite.org:32325"))
```

## List MetadataFormats

```{r}
list_metadataformats(id = "oai:oai.datacite.org:32348")
```

## List Sets

```{r}
list_sets("http://oai.datacite.org/oai")
```
