---
name: usda-plants-database-r
layout: post
title: USDA plants database API in R
date: 2016-10-19
author: Scott Chamberlain
sourceslug: _posts/2016-10-19-usda-plants-database-r.md
tags:
- API
- R
- ecology
- plants
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  collapse = TRUE
)

options(tibble.max_extra_cols = 15)
```

The USDA maintains a database of plant information, some of it trait data, some
of it life history. Check it out at <http://plants.usda.gov/java/>

They've been talking about releasing an API for a long time, but have not done so.

Thus, since at least some version of their data is in the public web,
I've created a RESTful API for the data:

* source code: <https://github.com/sckott/usdaplantsapi/>
* base URL: <https://plantsdb.xyz>

Check out the API, and open issues for bugs/feature requests in the github repo.

The following is an example using it from R, but you can use it from anywhere,
the command line, Ruby, Python, a browser, whatevs.

Here, we'll use [request](https://github.com/sckott/request), a higher level
http client for R that I've been working on. A small quirk with `request` is that
when piping, you have to assign the output of the request to an object before you
can do any further manipulation. But that's probably good for avoiding too long
pipe chains.

> note, that I've set `tibble.max_extra_cols = 15` to not print the many 
columns that are returned, for blog post brevity. When you run below
you'll get more columns.

Install from CRAN

```{r eval=FALSE}
install.packages("request")
```

There is a small improvement in the dev version of `request` to make any data.frame's
tibble's (which the below examples uses). To get that install from GitHub:

```{r eval=FALSE}
devtools::install_github("sckott/request")
```

```{r}
library('request')
library('tibble')
```

### Heartbeat

The simplest call to the API is to a route `/heartbeat`,
which just lists the available routes.

Set the base url we'll use throughout the work below

```{r}
root <- api("https://plantsdb.xyz")
```

```{r}
root %>% api_path(heartbeat)
```


Okay, so there are just two routes, `/search` and `/heartbeat`.

## Search

The search route suppports the following parameters:

- `fields`, e.g., `fields='Genus,Species'` (default: all fields returned)
- `limit`, e.g., `limit=10` (default: 10)
- `offset`, e.g., `offset=1` (default: 0)
- search on any fields in the output, e.g, `Genus=Pinus` or `Species=annua`

### basic search

Let's first not pass any params

```{r}
root %>% api_path(search)
```

You get slots:

* `count`: number of results found
* `returned`: numbef of results returned
* `citation`: suggested citation, from USDA
* `terms`: terms of use, from USDA
* `data`: the results
* `error`: if an error occurred, you'll see the message here

Note that if any data.frame's are found, we make them into tibble's, nicely
formatted data.frame's that make it easy to deal with large data.

### Pagination

limit number of results

```{r}
root %>%
  api_path(search) %>%
  api_query(limit = 5)
```

change record to start at

```{r}
root %>%
  api_path(search) %>%
  api_query(limit = 5, offset = 10)
```

### Return fields

You can say what fields you want returned, useful when you just want a
subset of fields

```{r}
root %>%
  api_path(search) %>%
  api_query(fields = 'Genus,Species,Symbol')
```


### Query

You can query on individual fields

```{r}
root %>%
  api_path(search) %>%
  api_query(Genus = Pinus, fields = "Genus,Species")
```

Another query example

```{r}
root %>%
  api_path(search) %>%
  api_query(Species = annua, fields = "Genus,Species")
```

And one more example, here we're interested in finding taxa that are perennials

```{r}
root %>%
  api_path(search) %>%
  api_query(Duration = Perennial, fields = "Genus,Species,Symbol,Duration")
```
