---
name: noaa-isd
layout: post
title: "NOAA ISD data in R"
date: 2019-10-07
author: Scott Chamberlain
sourceslug: _drafts/2019-10-07-noaa-isd.Rmd
tags:
- R
- NOAA
- climate
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.max_extra_cols = 15)
options(dplyr.max_extra_cols = 15)
```

## NOAA Integrated Surface Data (ISD)

NOAA hosts [Integrated Surface Data](https://www.ncdc.noaa.gov/isd), global
hourly data from ~35,000 stations worldwide (w/ best coverage in North America,
Europe, Australia, and parts of Asia), with data as far back as 1901; the data
sums up to around 600 GB!

## isdparser and rnoaa

[rnoaa][ghrnoaa] is a large-ish R package, with functions for accessing many different
data sources provided by NOAA.

One of those data sources is the ISD. All the internals for ISD data were at
one point inside rnoaa, but a while we broke off ISD parsing specifically (
that is, you've already downloaded an ISD data file, but only need to parse
the data in that file) into a new package called [isdparser][]. The format of 
data in ISD files is many lines like:

> 0076104270999991928040106004+51183+008483FM-12+025799999V020999990
> 04619999999N0000001N9+99999+99999999999ADDAY141061GF10899199905100
> 0251999999MD1310742+9999MW1451EQDQ01+000742APC3

Where spatial positions determine where a variable starts and ends, and then
each line also has optional data on the end of the line. It is definitely
not a fun job to parse this data; thus a package just to deal with these in
isdparser.

## Install packages

```{r eval=FALSE}
install.packages(c("isdparser", "rnoaa"))
```

```{r}
library(rnoaa)
library(isdparser)
```

## ISD in isdparser

isdparser assumes you already have or will get an ISD file yourself. 
There are of course example files you'd get from NOAA ISD in isdparser
package.

The isdparser package has three main functions:

- `isd_parse()` - Parse a NOAA ISD data file
- `isd_parse_line()` - Parse a single line of a NOAA ISD data file
- `isd_transform()` - Transform variables in output of `isd_parse()/isd_parse_line()`;
uses `isd_metadata`, a data.frame with metadata on units and scaling factors for
each variable

Read all lines in an ISD file

```{r}
library(isdparser)
path <- system.file('extdata/104270-99999-1928.gz', package = "isdparser")
isd_parse(path)
```

Read line by line

```{r}
path <- system.file('extdata/024130-99999-2016.gz', package = "isdparser")
lns <- readLines(path, encoding = "latin1")
isd_parse_line(lns[1])
```

With `isd_transform()` you can transform the data variables (including date,
latitude, longitude) to appropriate units.

```{r}
path <- system.file('extdata/104270-99999-1928.gz', package = "isdparser")
res <- isd_parse(path)
isd_transform(res)
```

You can also use `isd_transform()` on the output of `rnoaa::isd()`:

```{r}
x <- isd(104270, 99999, year = 1928)
head(x, n = 3)
head(isd_transform(x), n = 3)
```

## ISD in rnoaa

rnoaa has four functions for working with ISD data:

- `isd()` - Get ISD data for a given year and station, downloads data, then
uses `isdparser::isd_parse` to parse the data
- `isd_read()` - Read/parse an ISD data file (if already downloaded), thin wrapper
around `isdparser::isd_parse`. Good fit if you already have somem ISD files.
- `isd_stations()` - Fetch a data.frame of ISD stations and their metadata
- `isd_stations_search()` - Search for ISD stations by latitude, longitude,
radius, and bounding box

If you don't already have station identifiers in mind, you'll probably want to
start with searching for stations (searching from
a randomish spot on the coast in California, USA):

```{r}
library(rnoaa)
isd_stations_search(lat = 38.4, lon = -123, radius = 25)
```

Gives 7 stations. The Sonoma County Airport station seems like a good one to use,
with data from 2000 up to January 2019.

Then download data from that station with `isd()`

```{r}
isd(usaf = 724957, wban = 23213, year = 2008)
```

## To do

- ISD data is also [available in CSV format](https://github.com/ropensci/isdparser/issues/16)
as of last year. I haven't looked into it yet, but it's on the to do list


[ghrnoaa]: https://github.com/ropensci/rnoaa/
[crnoaa]: https://cloud.r-project.org/web/packages/rnoaa/
[ghisdp]: https://github.com/ropensci/isdparser/
[cisdp]: https://cloud.r-project.org/web/packages/isdparser/
[isdparser]: https://github.com/ropensci/isdparser/
