`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

*********
I had seen the function idata.frame in plyr before, but not really tested it. From the plyr documentation: 

> _"An immutable data frame works like an ordinary data frame, except that when you subset it, it returns a reference to the original data frame, not a a copy. This makes subsetting substantially faster and has a big impact when you are working with large datasets with many groups."_

For example, although baseball is a data.frame, its immutable counterpart is a reference to it:

```{r eval=FALSE}
> idata.frame(baseball)
<environment: 0x1022c74e8>
attr(,"class")
[1] "idf"         "environment"
```

Here are a few comparisons of operations on normal data frames and immutable data frames. Immutable data frames don't work with the doBy package, but do work with aggregate in base functions. &nbsp;Overall, the speed gains using idata.frame are quite impressive - I will use it more often for sure.

Here's the comparisons of idata.frames and data.frames:

*********

```{r}
# load packages
require(plyr); require(reshape2)

# Make immutable data frame
baseball_i <- idata.frame(baseball)
```

```{r}
# Example 1 - idata.frame more than twice as fast
system.time( replicate(50, ddply( baseball, "year", summarise, mean(rbi))) )
system.time( replicate(50, ddply( baseball_i, "year", summarise, mean(rbi))) )
```

```{r}
# Example 2 - Bummer, this does not work with idata.frame's
colwise(max, is.numeric) ( baseball ) # works
colwise(max, is.numeric) ( baseball_i ) # doesn't work
```

```{r}
# Example 3 - idata.frame twice as fast
system.time( replicate(100, baseball[baseball$year == "1884", ] ) )
system.time( replicate(100, baseball_i[baseball_i$year == "1884", ] ) )
```

```{r}
# Example 4 - idata.frame faster
system.time( replicate(50, melt(baseball[, 1:4], id = 1) ) )
system.time( replicate(50, melt(baseball_i[, 1:4], id = 1) ) )
```

```{r}
# And you can go back to a data frame by 
d <- as.data.frame(baseball_i)
str(d)
```

```{r}
# idata.frame doesn't work with the doBy package
require(doBy)
summaryBy(rbi ~ year, baseball_i, FUN=c(mean), na.rm=T)
```

```{r}
# But idata.frame works with aggregate in base (but with minimal speed gains)
# and aggregate is faster than ddply 
system.time( replicate(100, aggregate(rbi ~ year, baseball, mean) ) )
system.time( replicate(100, aggregate(rbi ~ year, baseball_i, mean) ) )
system.time( replicate(100, ddply( baseball_i, "year", summarise, mean(rbi)) ) )
```